---
title: "Read in network survey data"
---

Summary: Code to read in network survey data 

Notes:

* date_join_final_death = date of death 
* death_date_dc = date of interview 

```{r}
## library packages 
library(tidyverse)
library(here)

## read in respondent-level file  
results_ns <- readxl::read_xlsx(here("data/clean_data/2023-09-08/network_survey_2023-08-16.xlsx"), sheet = "main_networksurvival")

## read in death file  
deaths <- readxl::read_xlsx(here("data/clean_data/2023-09-08/network_survey_2023-08-16.xlsx"), sheet = "deaths")
```

```{r}
tmp <- deaths %>% select(starts_with('death_relationship'))
num_rel <- rowSums(tmp %>% 
                     mutate(`death_relationship/household` = as.numeric(`death_relationship/household`)) %>%
                     select(-death_relationship))

deaths_recoded <- deaths
deaths_recoded$num_rel <- num_rel
  
```


```{r}
results_ns %>% 
  mutate(today_date = as_date(today)) %>%
  ggplot(.) + geom_histogram(aes(x=today_date))
```




```{r}
## plot month of death distributions by month of report 
report_plot_df <- deaths_recoded %>% 
  mutate(death_month = lubridate::floor_date(x = as_date(date_join_final_death), unit = "month")) %>% 
  mutate(report_month = lubridate::floor_date(x = as_date(death_date_dc), unit = "month")) %>% 
  filter(report_month != as_date("2023-02-01")) %>% ## filter out February interviews  
  filter(!is.na(death_month)) 

report_plot <- report_plot_df %>% 
  ggplot(., aes(x = as.factor(death_month))) + 
    geom_bar(position=position_dodge2(.9), fill = "grey", color = "black") +
    geom_text(stat='count', aes(label=..count..), vjust = -.5) + 
      facet_wrap(~report_month, nrow = 4) + 
      ylim(0, 80) + 
      cowplot::theme_cowplot() + 
      labs(x = "Month of Death",
           title = "Reported Number of Deaths (Panels = Month of Interview)")

report_plot
```

### try joining respondent and death data

This dataset has deaths, with attributes of the respondent and interviewer appended

```{r}
deaths_resp <- deaths_recoded %>%
  left_join(results_ns %>% select(uuid_ki, start, end, today, enum:movement_other),
            by=c('uuid_ki4'='uuid_ki')) %>%
  mutate(today_date = as_date(today)) %>%
  mutate(today_interview_month = month(today_date))

deaths_resp
```

This dataset has respondents, with numbers of deaths added 

```{r}
resp_deaths <- results_ns %>%
  left_join(deaths_recoded %>%
              group_by(uuid_ki4) %>%
              #summarize(total_reported_deaths = sum(num_rel)), 
              count(name='total_reported_deaths'), 
            by=c('uuid_ki'='uuid_ki4')) %>%
  mutate(today_date = as_date(today)) %>%
  mutate(today_interview_month = month(today_date)) %>% 
  mutate(total_reported_deaths = ifelse(is.na(total_reported_deaths), 
                                        0, 
                                        total_reported_deaths))

resp_deaths
```

```{r}
resp_deaths %>%
  ggplot(.) +
  geom_point(aes(x = num_deaths_total, y = total_reported_deaths), alpha = .3) +
  theme_minimal()
```


```{r}
with(resp_deaths, table(total_reported_deaths, useNA='ifany'))
```

```{r}
with(deaths, table(death_pos, useNA='ifany'))
```


```{r}
with(resp_deaths, table(today_interview_month, useNA='ifany'))
```

Reported deaths by month

```{r}
resp_deaths %>%
  group_by(today_interview_month) %>%
  summarize(agg_reported_deaths = sum(total_reported_deaths),
            num_respondents = n()) %>%
  mutate(deaths_per_respondent = agg_reported_deaths / num_respondents)
```


```{r}
resp_deaths %>%
  group_by(today_interview_month) %>%
  summarize(avg_reported_deaths = mean(total_reported_deaths)) %>%
  ggplot(.) +
  geom_point(aes(x=today_interview_month, y=avg_reported_deaths)) +
  expand_limits(y=0) +
  theme_minimal()
```

```{r}
m1 <- lm(total_reported_deaths ~ today_interview_month*as.factor(enum),
         data = resp_deaths)
summary(m1)
```

```{r}
m2 <- lm(total_reported_deaths ~ as.factor(interview_site),
         data = resp_deaths)
summary(m2)
```

```{r}
m3 <- lm(total_reported_deaths ~ today_interview_month*as.factor(interview_site),
         data = resp_deaths)
summary(m3)
```


```{r}
resp_deaths %>%
  group_by(today_interview_month, interview_site) %>%
  summarize(avg_reported_deaths = mean(total_reported_deaths)) %>%
  ggplot(.) +
  geom_line(aes(x=today_interview_month, y=avg_reported_deaths, color=interview_site, group=interview_site)) +
  expand_limits(y=0) +
  theme_minimal()
```

 -

### Dig into characteristics of respondents

```{r}
m4 <- lm(total_reported_deaths ~ ki_sex*poly(ki_age, 2),
         data = resp_deaths %>%
           filter(! is.na(ki_age)))
summary(m4)
```

```{r}
m4 <- lm(total_reported_deaths ~ today_interview_month*ki_sex*poly(ki_age, 2),
         data = resp_deaths %>%
           filter(! is.na(ki_age)))
summary(m4)
```

Age/sex composition of respondents over time (totals)

```{r}
resp_deaths %>%
  mutate(ki_age_bin = cut_interval(ki_age, length=10)) %>%
  group_by(today_interview_month, ki_sex, ki_age_bin) %>%
  count(name='num_respondents') %>%
  ggplot(.) +
  geom_line(aes(x=today_interview_month, y=num_respondents, color=ki_sex, group=ki_sex)) +
  facet_wrap(~ ki_age_bin) +
  expand_limits(y=0) +
  theme_minimal()
```
Age/sex composition of respondents over time (prop in each time period)

```{r}
resp_deaths %>%
  mutate(ki_age_bin = cut_interval(ki_age, length=10)) %>%
  group_by(today_interview_month, ki_sex, ki_age_bin) %>%
  count(name='num_respondents') %>%
  ungroup() %>%
  group_by(today_interview_month) %>%
  mutate(prop_respondents = num_respondents/sum(num_respondents)) %>%
  ggplot(.) +
  geom_line(aes(x=today_interview_month, y=prop_respondents, color=ki_sex, group=ki_sex)) +
  facet_wrap(~ ki_age_bin) +
  expand_limits(y=0) +
  theme_minimal()
```

Total reported deaths over time by age/sex composition of respondents

```{r}
resp_deaths %>%
  mutate(ki_age_bin = cut_interval(ki_age, length=10)) %>%
  group_by(today_interview_month, ki_sex, ki_age_bin) %>%
  summarize(tot_reported_deaths = sum(total_reported_deaths)) %>%
  ggplot(.) +
  geom_line(aes(x=today_interview_month, y=tot_reported_deaths, color=ki_sex, group=ki_sex)) +
  facet_wrap(~ ki_age_bin) +
  expand_limits(y=0) +
  theme_minimal()
```
Avg reported deaths over time by age/sex composition of respondents

```{r}
resp_deaths %>%
  mutate(ki_age_bin = cut_interval(ki_age, length=10)) %>%
  group_by(today_interview_month, ki_sex, ki_age_bin) %>%
  summarize(avg_reported_deaths = mean(total_reported_deaths)) %>%
  ggplot(.) +
  geom_line(aes(x=today_interview_month, y=avg_reported_deaths, color=ki_sex, group=ki_sex)) +
  facet_wrap(~ ki_age_bin) +
  expand_limits(y=0) +
  theme_minimal()
```
### Deaths over time by relationship

Something is odd about these relationships -- some of these deaths have more than one

```{r}
tmp <- deaths %>% select(starts_with('death_relationship'))
num_rel <- rowSums(tmp %>% mutate(`death_relationship/household` = as.numeric(`death_relationship/household`)) %>% select(-death_relationship))

tmp_deaths <- deaths
tmp_deaths$num_rel <- num_rel
```




Avg reported deaths over time by relationship 

```{r}
deaths_resp %>%
  group_by(today_interview_month, death_relationship) %>%
  count(name='num_deaths') %>%
  left_join(resp_deaths %>%
              group_by(today_interview_month) %>%
              count(name='num_interviews')) %>%
  mutate(avg_reported_deaths = num_deaths / num_interviews) %>%
  ggplot(.) +
  geom_line(aes(x=today_interview_month, y=avg_reported_deaths, color=death_relationship, group=death_relationship)) +
  expand_limits(y=0) +
  theme_minimal()  
```

Avg reported deaths over time by relationship (proportion)

```{r}
deaths_resp %>%
  group_by(today_interview_month, death_relationship) %>%
  count(name='num_deaths') %>%
  left_join(resp_deaths %>%
              group_by(today_interview_month) %>%
              count(name='num_interviews')) %>%
  mutate(avg_reported_deaths = num_deaths / num_interviews) %>%
  ggplot(.) +
  geom_line(aes(x=today_interview_month, y=avg_reported_deaths, color=death_relationship, group=death_relationship)) +
  expand_limits(y=0) +
  theme_minimal()  
```
 
