---
title: "R Notebook"
output: html_notebook
---


## Basic Scale-up Estimator

To estimate under-5 mortality rate from the aggregate relational data, we can
apply the basic network scale-up estimator:

$$
\tag{2}
\widehat{M}_{\text{crude}} = \frac{\sum_{i \in s_F} w_i \enspace y_{i, H}}{\sum_{i \in s_F} w_i \enspace \hat{d}_{i}}
$$

where

-   $\widehat{M}_{\text{crude}}$ is the network survival estimate of the 6-month crude mortality rate 

-   $\hat{d}_{i}$ is the total degree (size of personal network)

-   $y_{i,H}$ is the total number of connections respondent $i$ has to people who have died in the last 6 months 

-   $s_F$ is the sampling frame

-   $w_i$ is the survey weight for respondent $i$


```{r}
library(tidyverse)
library(data.table)


pilot <- fread("~/workspace/reach/formative_data_networks_revised.csv")
```


```{r}
pilot %>% 
  ggplot() + 
  geom_histogram(aes(x = months_6), binwidth = 1, color = "black", fill = "grey") + 
  cowplot::theme_cowplot() + 
  labs(x = "Deaths Reported",
       title = "Count of Deaths (6 month recall)") 
```


```{r}
lists <- list()
  
for (i in 1:1000) {
  resample <- pilot %>% 
    sample_n(size = 500/3, replace = T) %>% 
    mutate(network = rnorm(n = n(), mean = 40, sd = 40)) %>% 
    summarize(num = sum(months_6),
              denom = sum(network)) %>% 
    mutate(estimate = num / denom)
  
  lists[[i]] <- resample
}

bind_rows(lists) %>%
  ggplot(aes(x = estimate)) + 
  geom_histogram(color = "black", fill = "grey") + 
  cowplot::theme_cowplot() + 
  xlim(0, .15)
  
bind_rows(lists) %>% 
  summarize(quantiles = quantile(estimate, probs = c(0.025, 0.975)))


```

