---
title: "Network Survival Estimation"
author: Casey Breen 
---

Summary: In this notebook, I provide code to estimate the network survival estimates of the mortality rate using the network survival method. Here, we use apply the basic network scale-up estimator; future work could consider more complex estimators.   


## Network Survival Estimator 

To estimate mortality rate from the aggregate relational data, we can apply the basic network scale-up estimator:

$$
\tag{2}
\widehat{M}_{\text{crude}} = \frac{\sum_{i \in s_F} w_i \enspace Y_{i, H}}{\sum_{i \in s_F} w_i \enspace \hat{d}_{i}}
$$

where

-   $\widehat{M}_{\text{crude}}$ is the network survival estimate of the crude mortality rate 

-   $\hat{d}_{i}$ is the total degree (size of personal network)

-   $y_{i,H}$ is the total number of connections respondent $i$ has to people who have died in the reference period

-   $s_F$ is the sampling frame

-   $w_i$ is the survey weight for respondent $i$


## Mortality rate adjustment

Before applying the above formula, we need to convert our estimates of mortality rates into proportions. For example, we can convert the following mortality (MR):

$$
\text{Mortality Rate (MR)} = \left(\frac{\text{Total Deaths During the Period}}{\text{Num. of people observed} \times  \text{Num. of observed days}}\right) \times 10,000
$$

into a proportion $p$ using the following equation:

$$ 
\begin{split}
P &= \left(\frac{\text{Num. Total Deaths During the Period}}{\text{Num. of people observed}} \right) \\ 
&= MR \times \frac{\text{Num. of observed days}}{10,000} 
\end{split} 
$$

This allows us to use the MDE equation above to estimate mortality. Here, we assume that the different length of reporting windows (e.g., 60 days vs. 180 days) are accounted for when we transform the mortality rate into a proportion of individuals who are during the reporting period.

# Code 

```{r}
## library packages 
library(tidyverse)
library(data.table)

## read in data 
results_ns <- readxl::read_xlsx("data/Tool1b_Clean_2023-03-02.xlsx", sheet = "main_networksurvival")
```

## Estimates for Kin Network Tie Definition 

Here, we apply the basic network survival estimator to the kin tie definition. 

```{r}
## estimates for kin tie definition  
results_ns %>% 
  summarize(total_kin_connections = sum(num_total_kin, na.rm = T), 
            n = sum(!is.na(num_total_kin)),
            total_kin_deaths = sum(num_deaths_kin, na.rm = T)) %>% 
  mutate(avg_kin_connections = total_kin_connections/n, 
         cmr_estimate = (total_kin_deaths/total_kin_connections)*(10000/80)) ## replace 80 with avg number of days since January 1st 
```

## Estimates for Neighbor Network Tie Definition 

Here, we apply the basic network survival estimator to the family tie definition 

```{r}

## estimates for neighbor tie definition 
results_ns %>% 
  summarize(total_neighbor_connections = sum(as.numeric(num_total_hh_neighbour), na.rm = T), 
            n = sum(!is.na(num_total_hh_neighbour)),
            total_neighbor_deaths = sum(num_deaths_neighbours, na.rm = T) + sum(num_deaths_hh, na.rm = T)) %>% 
    mutate(avg_neighbor_connections = total_neighbor_connections/n, 
           cmr_estimate = (total_neighbor_deaths/total_neighbor_connections)*(10000/80)) ## replace 80 with avg number of days since January 1st 


```

