---
title: "R Notebook"
author: Casey Breen
---


```{r}
## read in packages 
library(tidyverse)
library(data.table)
library(gt)
library(janitor)

## read in data 
results_ns <- readxl::read_xlsx("~/Downloads/Tool1b_Clean_2023-06-30.xlsx", sheet = "main_networksurvival")
deaths <- readxl::read_xlsx("~/Downloads/Tool1b_Clean_2023-06-30.xlsx", sheet = "deaths")
```


```{r}
results_ns_clean <- results_ns %>%
  filter(consent == "yes") 

results_ns_clean %>% 
  summarise(across(everything(), ~ sum(is.na(.x))))

```

Notes

* drop missing num_total_kin_male_child_under5 
* drop consent 
* drop num_total_kin
* num_deaths_male_child
* num_deaths_female_child
* num_deaths_male_grandchild
* num_deaths_female_grandchild

```{r}
results_ns_clean %>% 
  filter(is.na(num_total_hh_neighbour) | is.na(num_total_neighbour)) %>% 
  select(num_total_hh_neighbour, num_total_neighbour)

results_ns_clean %>% 
  filter(is.na(num_deaths_kin))

results_ns_clean %>% 
  filter(is.na(num_deaths_neighbours))

results_ns_clean %>% 
  filter(is.na(num_deaths_hh))
```


```{r}
results_ns_clean %>% 
  filter(is.na(ki_sex))

results_ns_clean %>% 
  filter(is.na(ki_age))

results_ns_clean %>% 
  select(material_house, education_level, marital_status, cooking_fuel, bed, radio, livelihood) %>% 
  summarise(across(everything(), ~ sum(is.na(.x))))
```


First, we drop any denominator (neighbors) with missing data (N = 3) and denominator (kin) with missing data (n = 18). 

Next, we will drop numerators with missing data. We will observations with both missing data for households (13), neighbors (2), and kin (20). 

Finally, we will drop data with missing weighting fields. Filtering out material house, education level, marital status, cooking fuel (1), bed (1), radio (1), and livelihood (1). 

## Clean Script 
```{r}
## Clean script

## restrict to consent == yes 
results_ns_clean <- results_ns %>%
  filter(consent == "yes") 

## drop missing neighbor, kin, and 
results_ns_clean <- results_ns_clean %>% 
  filter(!is.na(num_total_hh_neighbour) & !is.na(num_total_neighbour) & !is.na(num_deaths_kin))

## drop missing deaths 
results_ns_clean <- results_ns_clean %>% 
  filter(!is.na(num_deaths_kin) & !is.na(num_deaths_neighbours) & !is.na(num_deaths_hh))

## drop missing weighting variables  
results_ns_clean <- results_ns_clean %>% 
  filter(!is.na(cooking_fuel) & !is.na(bed) & !is.na(radio) & !is.na(livelihood))


```

```{r}
## deaths 
deaths <- deaths %>% 
  filter(!date_join_final_death %in% c(44927, 45061, 44928, 44951, 45030))
```





